---
title: "Introduction to RamEx"
output: rmarkdown::html_vignette
description: |
   A package for comprehensive and efficient analysis and comparison of ramanomes.
vignette: >
  %\VignetteIndexEntry{Introduction to RamEx}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Ramanome Explorer (RamEx, Fig.1) is a R package for comprehensive and efficient analysis and comparison of ramanomes.
By providing simple-to-use modules for computational tasks that otherwise would require substantial programming experience and algorithmic skills, RamEx should greatly facilitate the computational mining of ramanomes.

More complete information about how to use RamEx can be found in the [github](https://github.com/qibebt-bioinfo/RamEx), but here you'll find a brief overview of the function and examples to mining of ramanomes:

##Installaiton

```{r install_info, echo = FALSE}
#| RamEx will be installed from GitHub:
library('devtools');
install_github("qibebt-bioinfo/RamEx")
```

##Getting Started

```{r package_load, echo=FALSE}
#| after installed, you can load the package:
library('RamEx')
```

Once the package loaded, the example data RamEx-data can be loaded automatically. RamEx also accepts manual data loading.
```{r data_load, echo=FALSE}
#| after the package loaded, you can load the data manually:
RamEx-data <- read_spec("the folder path of your specture")
```
## Preprocessing

```{r data_preprocessing, echo=FALSE}
#| after the data loaded, data should preprocessed firstly:
RamEx-data <- pre.smooth.sg(data)
RamEx-data <- pre.baseline.polyfit(data)
RamEx-data <- pre.normalize.ch(data)
```
## Quality control:
```{r data quality control, echo=FALSE}
RamEx-data_index <- qc_jump(data@datasets$normalized.data,var_tol = 0.4)
RamEx-data <- RamEx-data[RamEx-data_index$index_good]
```

## Meta-based analysis
```{r meta-based analysis, echo=FALSE}
marker_info <- find_markers_roc(RamEx-data@datasets$normalized.data,RamEx-data@meta.data$group)
```

## Meta-free analysis
```{r meta-free analysis, echo=FALSE}
cluster_info <- louvain_clustering(object = RamEx-data, resolutions = c(0.01))
```


If you want to learn more, be sure to take a look at the [RamEx github page](https://github.com/qibebt-bioinfo/RamEx).

