---
title: "RamEx Vignette"
output: rmarkdown::html_vignette
description: |
   A package for comprehensive and efficient analysis and comparison of ramanomes.
vignette: >
  %\VignetteIndexEntry{RamEx Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Ramanome Explorer (RamEx, Fig.1) is a R package for comprehensive and efficient analysis and comparison of ramanomes.
By providing simple-to-use modules for computational tasks that otherwise would require substantial programming experience and algorithmic skills, RamEx should greatly facilitate the computational mining of ramanomes.

More complete information about how to use RamEx can be found in the [github](https://github.com/qibebt-bioinfo/RamEx), but here you'll find a brief overview of the function and examples to mining of ramanomes:

##Installaiton



##Getting Started
# Data Loading
Raman spectra are respectively tracked in single txt files, and their meta info is recorded in the file name.
Here we assume there's only one factor of the dataset, which means RamEx do not contain multiple-factor analysis. If you have multiple factors but they are independent of each other, these factors will be treated as one factor.
```{r}
library(RamEx)
library(magrittr)
data(RamEx_data)
data <- RamEx_data
options(mc.cores = 2)
```
# Pretreatment
Spectral pretreatment will make the spectrum clearer, containing smoothing, baseline removal, normalization and truncation.
Mean spectra will display their effects.
Here the results of each step will be kept in the Ramanome for better debugging, and 'draw.mean' exhibit the final dataset.
```{r}
data_spike <- Preprocesssing.Background.Spike(data, "CPU") 
data_smoothed <- Preprocesssing.Smooth.Sg(data) 
data_baseline <- Preprocesssing.Baseline.Polyfit(data_smoothed)
data_baseline_bubble <- Preprocesssing.Baseline.Bubble(data_smoothed)
data_normalized <- Preprocesssing.Normalize(data_baseline, "ch")
Preprocesssing.Cutoff(data_normalized,550, 1800)
mean.spec(data_normalized@datasets$baseline.data, data@meta.data$group) 
```
# Quality control
```{r}
data_cleaned <- Qualitycontrol.ICOD(data_normalized@datasets$normalized.data,var_tol = 0.4)
data_cleaned <- data_normalized[data_cleaned$index_good,] 
mean.spec(data_cleaned@datasets$normalized.data, data_cleaned@meta.data$group)
#qc_icod <- Qualitycontrol.Mcd(data_normalized@datasets$normalized.data)
#qc_t2 <- Qualitycontrol.T2(data_normalized@datasets$normalized.data) 
qc_dis <- Qualitycontrol.Dis(data_normalized@datasets$normalized.data)
hist(qc_dis$dis)
qc_snr <- Qualitycontrol.Snr(data_normalized@datasets$normalized.data) 
```

# Interested Bands
Get single-cell intensitiy or intensity accumulationy within a wavenumber range, pls give a list containing multiple bands or band ranges. These feature selection results will be saved as 'interested.bands' in the given Ramanome object. Further, you can add some equations by yourself. 
```{r}
data_cleaned <- Feature.Reduction.Intensity(data_cleaned, list(c(2000,2250),c(2750,3050), 1450, 1665))
# calculate CDR
CDR <- data.frame(data_cleaned@meta.data, 
                  data_cleaned@interested.bands$`2000~2250`/(data_cleaned@interested.bands$`2000~2250` + data_cleaned@interested.bands$`2750~3050`))
```

# Reduction
Nonlinear methods, such as UMAP and t-SNE. Linear methods like PCA, pCoA. The reduced sample matrix will be contained in the Ramanome onject as 'reductions'. Attention: RamEx uses PCA to reduce the dimensions of the high-dimensional spectrum, since UMAP and t-SNE are highly complex algorithms.
```{r}
data_cleaned <- Feature.Reduction.Pca(data_cleaned, draw=TRUE, save = FALSE) %>% Feature.Reduction.Pcoa(., draw=TRUE, save = FALSE) %>% Feature.Reduction.Tsne(., draw=TRUE, save = FALSE) %>% Feature.Reduction.Umap(., draw=TRUE, save=FALSE) #
``` 

# Markers analysis
```{r}
ROC_markers <- Raman.Markers.Roc(data_cleaned@datasets$normalized.data[,sample(1:1000, 50)],data_cleaned@meta.data$group) 
#cor_markers <- Raman.Markers.Correlations(data_cleaned@datasets$normalized.data[,sample(1:1000, 50)],as.numeric(data_cleaned@meta.data$group), min.cor = 0.6)
RBCS.markers <- Raman.Markers.Rbcs(data_cleaned, threshold = 0.003, draw = FALSE) 
```
# IRCA
-Global IRCA
```{r}
IRCA.interests <- Intraramanome.Analysis.Irca.Global(data_cleaned)
```
-Local IRCA
```{r}
bands_ann <- data.frame(rbind(cbind(c(742,850,872,971,997,1098,1293,1328,1426,1576),'Nucleic acid'),
                              cbind(c(824,883,1005,1033,1051,1237,1559,1651),'Protein'),
                              cbind(c(1076,1119,1370,2834,2866,2912),'Lipids')))
colnames(bands_ann) <- c('Wave_num', 'Group')
Intraramanome.Analysis.Irca.Local(data_cleaned, bands_ann = bands_ann)
```
- 2D-COS
```{r}
#Intraramanome.Analysis.2Dcos(data_cleaned) 
```
# Phenotype analysis
```{r}
#clusters_louvain <- Phenotype.Analysis.Louvaincluster(object = data_cleaned, resolutions = c(0.8)) 
clusters_kmneans <- Phenotype.Analysis.Kmeans(data_cleaned)
clusters_hca <- Phenotype.Analysis.Hca(data_cleaned) 
```

# Classifications
```{r}
Classification.Gmm(data_cleaned) 
Classification.Lda(data_cleaned)
Classification.Rf(data_cleaned)
Classification.Svm(data_cleaned)
```
# Quantifications
```{r}
quan_pls <- Quantification.Pls(data_cleaned)
quan_mlr <- Quantification.Mlr(data_cleaned)
quan_glm <- Quantification.Glm(data_cleaned)
```

# Spectral decomposition
```{r}
decom_mcr <- Spectral.Decomposition.Mcrals(data_cleaned,2)
decom_ica <- Spectral.Decomposition.Ica(data_cleaned, 2) 
data_nmf <- data_cleaned
data_nmf@datasets$normalized.data %<>% abs
decom_nmf <- Spectral.Decomposition.Nmf(data_nmf)
