---
title: "Introduction to RamEx"
output: rmarkdown::html_vignette
description: |
   A package for comprehensive and efficient analysis and comparison of ramanomes.
vignette: >
  %\VignetteIndexEntry{Introduction to RamEx}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Ramanome Explorer (RamEx, Fig.1) is a R package for comprehensive and efficient analysis and comparison of ramanomes.
By providing simple-to-use modules for computational tasks that otherwise would require substantial programming experience and algorithmic skills, RamEx should greatly facilitate the computational mining of ramanomes.

More complete information about how to use RamEx can be found in the [github](https://github.com/qibebt-bioinfo/RamEx), but here you'll find a brief overview of the function and examples to mining of ramanomes:

##Installaiton
```{r}
library(devtools)
install_github("qibebt-bioinfo/RamEx")
```
##Getting Started
# Data Loading
Raman spectra are respectively tracked in single txt files, and their meta info is recorded in the file name.
```{r}
library(RamEx)
library(dplyr)
data(RamEx_data)
options(mc.cores = 2)
```
# Pretreatment
SSpectral pretreatment will make the spectrum clearer, containing smoothing, baseline removal, normalization and truncation.
Mean spectra will display their effects.
Here the results of each step will be kept in the Ramanome for better debugging, and 'mean.spec' exhibit the final dataset.
```{r}
RamEx_data <- RamEx_data %>% Preprocessing.Smooth.Sg %>% Preprocessing.Baseline.Polyfit %>% Preprocessing.Normalize(.,'ch') 
mean.spec(RamEx_data$normalized.data, RamEx_data$group)  
```
# Quality control
Remove the samples with low quality or abnormal spectral acquisition. ICOD method is recommended to perform quality control (QC). Each method will return a logical vector, indicating the quality of each cell sample.
```{r}
qc_icod <- Qualitycontrol.ICOD(RamEx_data)
data_cleaned <- RamEx_data[qc_icod$quality,] 
mean.spec(data_cleaned$normalized.data, data_cleaned$group,0.3)
qc_mcd <- Qualitycontrol.Mcd(RamEx_data) 
qc_t2 <- Qualitycontrol.T2(RamEx_data) 
qc_dis <- Qualitycontrol.Dis(RamEx_data) 
qc_snr <- Qualitycontrol.Snr(RamEx_data, 'easy')
```

# Interested Bands
Get single-cell intensitiy or intensity accumulationy within a wavenumber range, pls give a list containing multiple bands or band ranges. These feature selection results will be saved as 'interested.bands' in the given Ramanome object. Further, you can add some equations by yourself (e.g. calculate CDR). 
```{r}
data_cleaned <- Feature.Reduction.Intensity(data_cleaned, list(c(2000,2250),c(2750,3050), 1450, 1665))
# calculate CDR
CDR <- data.frame(data_cleaned@meta.data, 
                  data_cleaned$`2000~2250`/(data_cleaned$`2000~2250` + data_cleaned$`2750~3050`))
```

# Reduction
Nonlinear methods, such as UMAP and t-SNE. Linear methods like PCA, pCoA. The reduced sample matrix will be contained in the Ramanome onject as 'reductions'. Attention: A pre-treatment of large spectral matrix by PCA is recommended, since UMAP and t-SNE are highly complex algorithms.
```{r}
data.reduction <- Feature.Reduction.Pca(data_cleaned) %>% Feature.Reduction.Pcoa %>% Feature.Reduction.Tsne %>% Feature.Reduction.Umap
``` 

# Markers analysis
RamEx provides three types of markers analysis: ROC, correlation and RBCS. 
In addition to singular Raman markers, paired bands are also examined
```{r}
ROC_markers <- Raman.Markers.Roc(Preprocessing.Cutoff(data_cleaned, 1400, 1500), paired  = TRUE, threshold = 0.6) 
cor_markers <- Raman.Markers.Correlations(Preprocessing.Cutoff(data_cleaned, 1400, 1500), min.cor = 0.6) 
RBCS.markers <- Raman.Markers.Rbcs(data_cleaned, threshold = 0.003, show = T) 
```
# IRCA
Observe the single-cell heterogeneity within the Ramanome/microbial community and to elucidate intracellular metabolite conversions.
- Global IRCA.
```{r}
IRCA.interests <- Intraramanome.Analysis.Irca.Global(data_cleaned)
```
- Local IRCA
```{r}
bands_ann <- data.frame(
  Wave_num = c(742,850,872,971,997,1098,1293,1328,1426,1576,
               824,883,1005,1033,1051,1237,1559,1651,
               1076,1119,1370,2834,2866,2912),
  Group = rep(c('Nucleic acid', 'Protein', 'Lipids'), times = c(10, 8, 6)) )
Intraramanome.Analysis.Irca.Local(data_cleaned, bands_ann = bands_ann)
```
- 2D-COS
Captures both synchronous (simultaneous changes) and asynchronous (sequential changes) relationships, providing detailed insights into spectral changes
```{r}
data_cos <- Intraramanome.Analysis.2Dcos(data_cleaned) 
```
# Phenotype analysis
RamEx provides three types of phenotype analysis: Louvain clustering, K-means clustering and hierarchical clustering.
```{r}
clusters_louvain <- Phenotype.Analysis.Louvaincluster(object = data_cleaned, resolutions = c(0.8)) 
clusters_kmneans <- Phenotype.Analysis.Kmeans(data_cleaned,5)
clusters_hca <- Phenotype.Analysis.Hca(data_cleaned)
```

# Classifications
Gaussian Mixture Model (GMM), Linear Discriminant Analysis (LDA), Random Forest (RF), Support Vector Machine (SVM)
```{r}
model.gmm <- Classification.Gmm(data_cleaned)
model.lda <- Classification.Lda(data_cleaned)
model.rf <- Classification.Rf(data_cleaned)
model.svm <- Classification.Svm(data_cleaned)
# when new data is available, you can use the following code to predict the new data by using the trained model.
pred_new <- predict_classification(model.lda, data_cleaned)
```
# Quantifications
Partial Least Squares (PLS), Multiple Linear Regression (MLR), Generalized Linear Model (GLM)
```{r}
quan_pls <- Quantification.Pls(data_cleaned) 
quan_mlr <- Quantification.Mlr(data_cleaned) 
quan_glm <- Quantification.Glm(data_cleaned) 
# when new data is available, you can use the following code to predict the new data by using the trained model.
pred_new <- predict_quantification(quan_pls, data_cleaned)
```

# Spectral decomposition
Multiple Component Regression - Alternating Least Squares (MCR-ALS), Independent Component Analysis (ICA), Non-negative Matrix Factorization (NMF) are provide to decompose the spectral matrix into multiple components.
```{r}
decom_mcr <- Spectral.Decomposition.Mcrals(data_cleaned,n_comp = 2)
decom_ica <- Spectral.Decomposition.Ica(data_cleaned, n_comp = 2) 
decom_nmf <- Spectral.Decomposition.Nmf(data_cleaned, n_comp = 2) 
```